Title Goes Here

### 2017-08-21 analysis/paper/paper.Rmd
stash

---
title: "A Standard for the Scholarly Citation of Archaeological Data: Supplementary Information"
author:
  - "Ben Marwick"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
    fig_caption: yes
    reference_docx: templates/template.docx
bibliography: references.bib
csl: journal-of-archaeological-science.csl
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

This document contains code and text supporting the results presented in "A Standard for the Scholarly Citation of Archaeological Data" (draft online at https://docs.google.com/document/d/1UT4RuW4SCkdLGsn2UmNK22vgfJZxljg8ugRstjToCN8/edit).

The source file for this document, and other files in the compendium are online at https://github.com/benmarwick/datacitation

@@

In addition to our email survey, we also conducted a pilot study of a random sample of `r nrow(jas_article_survey)` articles published during 2016-2017 in the _Journal of Archaeological Science_ to investigate data sharing behaviours. In this sample we found openly available raw data in `r data_are_avail` papers (`r data_are_avail_perc`%), even though only `r have_data_avail_statement` articles (`r have_data_avail_statement_perc`%) include a data availability statement. For papers where data was available, `r data_in_supp` papers (`r data_in_supp_perc`%) have data contained in supplementary files published with the paper, `r data_in_table` (`r data_in_table_perc`%) have all the raw data in tables in the text of the paper, and `r data_in_repo` (`r data_in_repo_perc`%) have data in an online repository. The most frequently shared type of data is compositional (e.g. element concentrations, n = `r comp_n`, `r comp_perc`%), followed by radiocarbon and luminescence age data (where the dates are the primary object of analysis, n = `r dates_n`, `r dates_perc`%) and DNA sequences (n = `r dna_n`, `r dna_perc`%). Generally, compositional data is presented in tables in the text, dating data is presented in supplementary files, and DNA data are in repositories (as is typical for DNA data). Data shared in supplementary files and online repositories were most often in Microsoft Excel format (n = `r xlsx_n`, `r xlsx_perc`% of the files in supplements files and repositories), followed by CSV (comma separated variables, a open source plain text spreadsheet format, n = `r csv_n`, `r csv_perc`%), and tables embedded in Microsoft Word or PDF documents (n = `r doc_n`, `r doc_perc`%). When a scripting language such as R, Python, MATLAB or OxCal was used as the analysis software for a paper (n = `r script_n`, `r script_perc`% of the papers where would identify the software), only 2 papers also provided the script with the paper.

@@

```{r eval = TRUE}
# install.packages("rdatacite")

# Original data aquisition on 14 Aug 2017
# library(rdatacite)
# dc_arch <- dc_search(q = "archaeology", rows = pretty(100000)[2])

@@

```{r eval = TRUE}
library(tidyverse)
library(stringr)
# glimpse(dc_arch)

dc_arch <- 
dc_arch %>% 
mutate(the_text =  toupper(gsub(" |_|\\(|\\)", "-", rights)),
       the_text = gsub("(\\-)\\1+", "\\1", the_text),
       the_text = gsub("-$", "", the_text)) %>% 
  mutate(cc = case_when(
    grepl("SEMANTICS/OPENACCESS", the_text) ~ "EU Open Access",
    grepl("SEMANTICS/RESTRICTEDACCESS", the_text) ~ "EU Restricted Access",
    grepl("SEMANTICS/CLOSEDACCESS", the_text) ~ "EU Closed Access",
    grepl("CC-BY-[^NC-ND]|CC-BY$|ATTRIBUTION-LICENSE|ATTRIBUTION-[^NON]", the_text) ~ "CC-BY",
    grepl("CC-BY-NC-[^ND]|BY-NC-[^ND]|CC-NC$|ATTRIBUTION-NON-COMMERCIAL-", the_text) ~ "CC-BY-NC",
    grepl("CC-BY-NC-SA-[^ND]|CC-NC-SA$|ATTRIBUTION-NON-COMMERCIAL-SHARE", the_text) ~ "CC-BY-NC-SA",
    grepl("CC-BY-NC-ND|BY-NC-ND|ATTRIBUTION-NONCOMMERCIAL-NO-DERIVATIVE|ATTRIBUTION-NONCOMMERCIAL-NODERIVATIVES|ATTRIBUTION-NONCOMMERCIAL-NODERIVS|NAMENSNENNUNG-NICHTKOMMERZIELL-KEINEBEARBEITUNG|NAMENSNENNUNG-KEINE-KOMMERZIELLE-NUTZUNG", the_text) ~ "CC-BY-NC-ND",
    grepl("ADS|ARCHAEOLOGYDATASERVICE", the_text) ~ "ADS",
    grepl("CC0|CC-0|ZERO", the_text) ~ "CC-0"
  )) %>% 
  mutate(yyyy_date = as.numeric(substr(date, 1,4)),
         yyyy_created = as.numeric(substr(created, 1,4))) %>% 
  mutate(yyyy_date = if_else(is.na(yyyy_date), as.numeric(publicationYear), yyyy_date)) %>% 
  filter(yyyy_date > 1500) %>% 
  filter(yyyy_date < 2018) %>% 
  filter(!grepl("PANGAEA|Global Biodiversity", datacentre)) # pesky Mediterranean Data Archaeology and Rescue & Danish plants!
```

@@

```{r first-dates, eval = TRUE}
dates <- 
  dc_arch

@@

# when is the break out
dates_tally <- 
  dates %>% 
  group_by(yyyy_date) %>% 
  tally()

# when is that big peak, it's 2009
library(changepoint)
out <- cpt.meanvar(dates_tally$n, test.stat = "Poisson", method = "BinSeg")
breakouts <- dates_tally[cpts.ts(out), ]

# who is the top data source in each year?
top_data_sources <- 
dates %>% 
  group_by(datacentre) %>% 
  tally(sort = TRUE) %>% 
  slice(1:10) %>% 
  pull(datacentre)

top_data_source_each_year <- 
dates %>% 
  group_by(datacentre, yyyy_date) %>% 
  tally() %>%  
  ungroup(datacentre) %>% 
  group_by(yyyy_date) %>% 
  slice(which.max(n)) %>% 
  arrange(desc(yyyy_date))

top_data_sources <- 
  unique(top_data_source_each_year$datacentre)

 years_of_interest <- c(2009)
 ymax <- 10000

date_plot <- 
  dates  %>% 
  filter(!is.na(yyyy_date)) %>% 
  ggplot(aes(yyyy_date,
             fill = datacentre
             )) +
  geom_bar() +
  # guides(fill=FALSE) +
  scale_fill_viridis(discrete = TRUE,
                     breaks = top_data_sources[1:4],
                     labels = top_data_sources[1:4]) +
  theme_new() +
  theme(legend.position = c(0.3, 0.75)) +
  xlim(c(1950,2017)) +
  xlab("Year data was collected") +
  ylab("Number of archaeology DOIs\nknown to DataCite") +
  geom_vline(xintercept = years_of_interest, colour = "red") +
  annotate("text", 
           x = years_of_interest + 2,
           y = ymax,
           label = years_of_interest)

# plotly::ggplotly(date_plot2) %>%
#   layout(showlegend = FALSE)

@@

```{r file-types, eval = TRUE}
# File types

@@

Our third investigation of archaeologists' data sharing behaviours is an analysis of all records (n = `r prettyNum(nrow(dc_arch), big.mark = ",")`, as of August 2017) of archaeological data files in online repositories tracked by DataCite (an organisation that provides persistent identifiers, or DOIs for research data). The majority of records come from the Archaeology Data Service (ADS) in the UK, followed by the Dutch Archiving and Network Services (DANS-EASY). Although the earliest data dates to `r first_date`, most of the data was collected after 1957. The was a steady increase in the amount of data deposited during 1990-2008, as personal computers become more affordable and data easier to digitize. There is a substantial increase in 2009 when the DANS-EASY

Many of these records are grey literature such as unpublished reports, produced by commercial consultants or research projects. Of  the `r prettyNum(number_of_records_with_file_type, big.mark = ",")` (`r perc_of_records_with_file_type`%) records that include infomation about the file types in the deposit, `r perc_only_pdf`% (n = `r prettyNum(n_only_pdf, big.mark = ",")`) contain only PDFs. Structured data files such as spreadsheets, databases or shapefiles are found either alone or together with other file types (typically PDFs) in `r perc_structured_data_file`% (n = `r prettyNum(n_structured_data_file, big.mark = ",")`) records.

```{r eval = TRUE}
datacentre_plot <- 
  datacentre %>% 
  mutate(Var1 = str_wrap(Var1, 20)) %>% 
  slice(1:10) %>% 
ggplot(aes(reorder(Var1, n),
           n)) +
  geom_col() +
  coord_flip() +
  xlab("Data centre") +
  ylab("") +
  theme_new() +
  theme(axis.text.y  = element_text(size = 8))

@@

# show plots
datacentre_plot
allocator_plot
creator_plot
language_plot
contributor_plot
rights_plot
publisher_plot

@@

```{r dates, eval = TRUE}

# why so many in 2009? http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0078080 it's the Dutch

top_per_year <- 
dates %>% 
  group_by(yyyy_date, datacentre) %>% 
   tally(sort = TRUE)  %>% 
   top_n(n=1)

why_2013 <- 
  xx %>% 
  filter(as.numeric(substr(date, 1,4)) == 2013)  %>% 
  group_by(datacentre) %>% 
  tally(sort = TRUE)

why_2012 <- 
  xx %>% 
  filter(as.numeric(substr(date, 1,4)) == 2012)  %>% 
  group_by(datacentre) %>% 
  tally(sort = TRUE)

why_2011 <- 
  xx %>% 
  filter(as.numeric(substr(date, 1,4)) == 2011)  %>% 
  group_by(datacentre) %>% 
  tally(sort = TRUE)

why_2010 <- 
xx %>% 
  filter(as.numeric(substr(date, 1,4)) == 2010)  %>% 
  group_by(datacentre) %>% 
  tally(sort = TRUE)

why_2009 <- 
  xx %>% 
  filter(as.numeric(substr(date, 1,4)) == 2009)  %>% 
  group_by(datacentre) %>% 
  tally(sort = TRUE)

why_2008 <- 
  xx %>% 
  filter(as.numeric(substr(date, 1,4)) == 2008)  %>% 
  group_by(datacentre) %>% 
  tally(sort = TRUE)

# in 2009, among the Dutch, what is going on?
dutch_2009 <- 
dc_arch %>% 
  filter(yyyy_date == 2009,
         datacentre == "DELFT.EASY - DANS-EASY") %>% 
  group_by(publisher) %>% 
  tally(sort=TRUE)

# how about in ADS?
ads_2009 <- 
dc_arch %>% 
  filter(yyyy_date == 2009,
         grepl("ADS", datacentre)) %>% 
  group_by(creator) %>% 
  tally(sort=TRUE)

# compare to 2009
dutch_2008 <- 
dc_arch %>% 
  filter(yyyy_date == 2008,
         datacentre == "DELFT.EASY - DANS-EASY") %>% 
  group_by(publisher) %>% 
  tally(sort=TRUE) 
```

```{r}
# panel of plots
# combine plots for lang and data center
datacentre_and_langugage_plot <- 
  datacentre_plot + annotation_custom(ggplotGrob(language_plot),
                                      xmin = 1.15, 
                                      xmax = 6.15,
                                      ymin = 10000,
                                      ymax = 40000
  )

@@

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE, eval = TRUE}
# which R packages and versions?
devtools::session_info()
```

The current Git commit details are:

```{r eval = TRUE}
# what commit is this file at? You may need to change the path value
# if your Rmd is not in analysis/paper/
git2r::repository("../..")
```

>>>>

---
title: "Title Goes Here"
author:
  - author 1
  - author 2
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::html_document2:
    fig_caption: yes
    reference_docx: templates/template.docx
bibliography: references.bib
csl: journal-of-archaeological-science.csl
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

@@

In addition to our email survey, we also conducted a pilot study of a random sample of `r nrow(jas_article_survey)` articles published during 2016-2017 in the _Journal of Archaeological Science_ to investigate data sharing behaviours. In this sample we found openly available raw data in `r data_are_avail` papers (`r data_are_avail_perc`%), even though only `r have_data_avail_statement` articles (`r have_data_avail_statement_perc`%) include a data availability statement. For papers where data was available, `r data_in_supp` papers (`r data_in_supp_perc`%) have data contained in supplementary files published with the paper, `r data_in_table` (`r data_in_table_perc`%) have all the raw data in tables in the text of the paper, and `r data_in_repo` (`r data_in_repo_perc`%) have data in an online repository. The most frequently shared type of data is compositional (e.g. element concentrations, n = `r comp_n`, `r comp_perc`%), followed by radiocarbon and luminescence age data (where the dates are the primary object of analysis, n = `r dates_n`, `r dates_perc`%) and DNA sequences (n = `r dna_n`, `r dna_perc`%). Data shared in supplementary files and online repositories were most often in Microsoft Excel format (n = `r xlsx_n`, `r xlsx_perc`% of the files in supplements files and repositories), followed by CSV (comma separated variables, a open source plain text spreadsheet format, n = `r csv_n`, `r csv_perc`%), and tables embedded in Microsoft Word or PDF documents (n = `r doc_n`, `r doc_perc`%). When a scripting language such as R, Python, MATLAB or OxCal was used as the analysis software for a paper (n = `r script_n`, `r script_perc`% of the papers where would identify the software), only 2 papers also provided the script with the paper.

@@

```{r get-citation-data, eval = TRUE}
# install.packages("rdatacite")

# Originat data aquisition on 14 Aug 2017
# library(rdatacite)
# dc_arch <- dc_search(q = "archaeology", rows = pretty(100000)[2])

@@

```{r eval = TRUE}

library(tidyverse)
library(stringr)
# glimpse(dc_arch)

dc_arch <- 
dc_arch %>% 
mutate(the_text =  toupper(gsub(" |_|\\(|\\)", "-", rights)),
       the_text = gsub("(\\-)\\1+", "\\1", the_text),
       the_text = gsub("-$", "", the_text)) %>% 
  mutate(cc = case_when(
    grepl("SEMANTICS/OPENACCESS", the_text) ~ "EU Open Access",
    grepl("SEMANTICS/RESTRICTEDACCESS", the_text) ~ "EU Restricted Access",
    grepl("SEMANTICS/CLOSEDACCESS", the_text) ~ "EU Closed Access",
    grepl("CC-BY-[^NC-ND]|CC-BY$|ATTRIBUTION-LICENSE|ATTRIBUTION-[^NON]", the_text) ~ "CC-BY",
    grepl("CC-BY-NC-[^ND]|BY-NC-[^ND]|CC-NC$|ATTRIBUTION-NON-COMMERCIAL-", the_text) ~ "CC-BY-NC",
    grepl("CC-BY-NC-SA-[^ND]|CC-NC-SA$|ATTRIBUTION-NON-COMMERCIAL-SHARE", the_text) ~ "CC-BY-NC-SA",
    grepl("CC-BY-NC-ND|BY-NC-ND|ATTRIBUTION-NONCOMMERCIAL-NO-DERIVATIVE|ATTRIBUTION-NONCOMMERCIAL-NODERIVATIVES|ATTRIBUTION-NONCOMMERCIAL-NODERIVS|NAMENSNENNUNG-NICHTKOMMERZIELL-KEINEBEARBEITUNG|NAMENSNENNUNG-KEINE-KOMMERZIELLE-NUTZUNG", the_text) ~ "CC-BY-NC-ND",
    grepl("ADS|ARCHAEOLOGYDATASERVICE", the_text) ~ "ADS",
    grepl("CC0|CC-0|ZERO", the_text) ~ "CC-0"
  )) %>% 
  mutate(yyyy_created =  as.numeric(substr(created, 1,4)),
         yyyy_date =  as.numeric(substr(date, 1,4)))
```

@@

```{r first-dates}
dates <- 
  dc_arch %>% 
  mutate(yyyy_date = as.numeric(substr(date, 1,4)),
         yyyy_created = as.numeric(substr(created, 1,4)))

@@

# when in that big peak
dates_tally <- 
  dates %>% 
  group_by(yyyy_created) %>% 
  tally(sort = TRUE)
```

@@

```{r}
# File types

@@

Our third investigation of archaeologists' data sharing behaviours is an analysis of all records (n = `r prettyNum(nrow(dc_arch), big.mark = ",")`, as of August 2017) of archaeological data files in online repositories tracked by DataCite (an organisation that provides persistent identifiers, or DOIs for research data). The majority of records come from the Archaeology Data Service (ADS) in the UK, followed by the Dutch Archiving and Network Services (DANS-EASY). Although the earliest data dates to `r first_date`, the first records in these repositories were created in `r first_created`. ...the big peak...

Many of these records are grey literature such as unpublished reports, produced by commercial consultants or research projects. Of  the `r prettyNum(number_of_records_with_file_type,  big.mark = ",")` (`r perc_of_records_with_file_type`%) recrods that include infomation about the file types in the deposit, `r perc_only_pdf`% (n = `r n_only_pdf`) contain only PDFs, followed by `r perc_only_image`% (n = `r prettyNum(n_only_image, big.mark = ",")`) that contain only image files. Structured data files such as spreadsheets, databases and shapefiles are found either alone or together with other file types (typically PDFs) in `r perc_structured_data_file`% (n = `r prettyNum(n_structured_data_file, big.mark = ","`) records.

```{r}
datacentre_plot <- 
  datacentre %>% 
  mutate(Var1 = str_wrap(Var1, 20)) %>% 
  slice(1:10) %>% 
ggplot(aes(reorder(Var1, n),
           n)) +
  geom_col() +
  coord_flip() +
  xlab("Data centre") +
  ylab("") +
  theme_new() +
  theme(axis.text.y  = element_text(size = 8))

@@

date_plot <- 
  dates %>% 
  mutate(yyyy = as.numeric(substr(Var1, 1,4))) %>% 
  select(yyyy) %>% 
  ggplot(aes(yyyy)) +
  geom_histogram() +
  theme_new() +
  xlim(c(1950,2017)) +
  xlab("Year") +
  ylab("")

library(plotly)
# ggplotly(date_plot)

format_plot <-

@@

```{r dates eval = TRUE}

# why so many in 2010? http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0078080 it's the Dutch

what_2010 <- 
  xx %>% 
  filter(as.numeric(substr(date, 1,4)) == 2010)

why_2013 <- 
  xx %>% 
  filter(as.numeric(substr(date, 1,4)) == 2013) %>% 
  group_by(publisher) %>% 
  tally(sort = TRUE)

why_2012 <- 
  xx %>% 
  filter(as.numeric(substr(date, 1,4)) == 2012) %>% 
  group_by(publisher) %>% 
  tally(sort = TRUE)

why_2011 <- 
  xx %>% 
  filter(as.numeric(substr(date, 1,4)) == 2011) %>% 
  group_by(publisher) %>% 
  tally(sort = TRUE)

why_2010 <- 
xx %>% 
  filter(as.numeric(substr(date, 1,4)) == 2010) %>% 
  group_by(publisher) %>% 
  tally(sort = TRUE)

why_2009 <- 
  xx %>% 
  filter(as.numeric(substr(date, 1,4)) == 2009) %>% 
  group_by(publisher) %>% 
  tally(sort = TRUE)

why_2008 <- 
  xx %>% 
  filter(as.numeric(substr(date, 1,4)) == 2008) %>% 
  group_by(publisher) %>% 
  tally(sort = TRUE)
```

```{r eval=TRUE}
# panel of plots
# combine plots for lang and data center
datacentre_and_langugage_plot <- 
  datacentre_plot + annotation_custom(ggplotGrob(language_plot),
                                      xmin = 1.15, 
                                      xmax = 6.15,
                                      ymin = 10000,
                                      ymax = 40000
  )

@@
